---
title: "Final coursework - Data Exploration and Visualisation"
author: |
  STAT5228_GroupB:  
  Task 1 : Meerja Hassanah B.N [2520427]  
  Task 2 : Golamaully Nasrin Razia [2523258]  
  Task 3 : Laurent Adriana Marie Bennette [2520333]  
  Task 4 : Suntoo Tushti [2520957]

date: "2026-04-22"
output: pdf_document

---
\newpage
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(moments)
library(dplyr)
library(GGally)
library(knitr)
library(kableExtra)
```

\newpage
**Overview**


This report presents an analysis of the data with the aim of identifying key business insights related to sales performance, customer behaviour, and operational efficiency. The dataset was first cleaned and validated to ensure accuracy, after which exploratory and focused analyses were conducted.

A specific focus was placed on the Moka district and the Electronics product category to enable deeper insights into high-value transactions and operational patterns.

---
**Task 1 : Data audit, tidy data and cleaning**

For Task 1, we have loaded the csv file in R software and has conducted the data audit, data tidying and cleaning. It is saved as cleaned_data.csv . We then filtered by district and product category and saved the focus data as Moka_data.csv. 

```{r}
##changing working directory
setwd("C:/Users/meerj/OneDrive/Desktop/Data Exploration and Visualisation/Final Assignment")

##TASK1: 
data<-read.csv("STAT5228_Portfolio_Data.csv",header=T,na.strings=c(" ","NA"))
View(data)
attach(data)
str(data)

summary(data)


###checking for any missingness of values 
colSums(is.na(data)) %>% sort(decreasing = TRUE)

###Variable 1: Transaction Id - checking for any duplications
duplicates <- data %>%
  count(transaction_id) %>%
  filter(n > 1)

nrow(duplicates)
str(data)

#from the above generated code, we have 50 duplications of transactions id

##removing of duplications
data <- data %>%
  distinct(transaction_id, .keep_all = TRUE)

#From the first summary we had 24000 observations for transaction ids
#After removing the duplications, we are left with 23950 - successfully remove the 50 duplications

####Variable 2: discount_pct
#we have 480 missing dicount percentage 
##missing discount to be set to 0
data <- data %>%
  mutate(discount_pct = replace_na(discount_pct, 0))

##verifying dicount limit 0-0.35

data %>% filter(discount_pct < 0 | discount_pct > 0.35)
data <- data %>% filter(discount_pct >= 0, discount_pct <= 0.35)

###Variable 3 : quantity, unit price MUR
data %>% filter(quantity <= 0 | unit_price_mur <= 0)
data <- data %>% filter(quantity > 0, unit_price_mur > 0)

##removed invalid values

###variable 4: gross_sales_mur, net_sales_mur, cost_mur, gross_profit_mur 

###addressing inconsistencies
##net sales = gross * (1-discount)

data <- data %>%
  mutate(calc_net_sales = gross_sales_mur * (1 - discount_pct),
         diff = abs(calc_net_sales - net_sales_mur))

summary(data$diff)
 ##rounding off the small differences
data <- data %>%
  mutate(net_sales_mur = calc_net_sales)

summary(data)

###variable 5: delivery times : in store
data <- data %>%
  mutate(delivery_minutes = ifelse(channel == "In-store", NA, delivery_minutes))


###variable 6: net sales mur
##checking for outliers
quantile(data$net_sales_mur, 0.99, na.rm = TRUE)
upper <- quantile(data$net_sales_mur, 0.99, na.rm = TRUE)
upper

plot(data$net_sales_mur)
data <- data %>% filter(net_sales_mur < upper)
plot(data$net_sales_mur) 
###removed the extra 1% outliers on top


###Variable 7 : date
data <- data %>%
  mutate(datetime = ymd_hms(datetime))
data <- data %>% filter(!is.na(datetime))

###variable 8 : satisfcation score 
sum(is.na(data$satisfaction_score))
## we have 2384 missing values 
mean(is.na(data$satisfaction_score)) * 100
## 9.991 % of the data is missing

##check where the missingness is coming from
##is it from delivery, in store or online??

data %>%
  group_by(channel) %>%
  summarise(missing_rate = mean(is.na(satisfaction_score)) * 100)


##visuals for satisfaction score
ggplot(data, aes(x = channel, fill = is.na(satisfaction_score))) +
  geom_bar(position = "fill") +
  labs(title = "Missing Satisfaction Score by Channel",
       y = "Proportion",
       fill = "Missing Satisfaction Score?")
###from the graph we can see that more or less the missingness is similar across all channels
### this may implies a random non response as not all customers are surveyed

###variable 9:  store_lat, store_lon 

missing_coords <- data %>%
  filter(is.na(store_lat) | is.na(store_lon))
nrow(missing_coords)

###which store is affected
missing_coords %>%
  count(store_name, district, sort = TRUE)

##remove only for spacial
data <- data %>%
  filter(!is.na(store_lat) & !is.na(store_lon))


###change format of date and datetime 
data <- data %>%
  mutate(date = as.Date(date))

data <- data %>%
  mutate(datetime = ymd_hms(datetime))
##check if it has worked
class(data$date)
class(data$datetime)
######classification of class
data$channel<-as.factor(data$channel)
data$customer_gender<-as.factor(data$customer_gender)
data$payment_method<-as.factor(data$payment_method)
data$product_subcategory<-as.factor(data
$product_subcategory)
data$customer_segment<-as.factor(data$customer_segment)
data$age_band<-as.factor(data$age_band)
data$product_category<-as.factor(data$product_category)
data$campaign<-as.factor(data$campaign)

nrow(data)

###save the MASTERDATASET as cleaned_data
write_csv(data, "cleaned_data.csv")
data<-read.csv("cleaned_data.csv",header=T)


###filtering data to our focus district/category
###focus district=moka
##focus category = electronics

focus_data<-data %>%
  filter(district == "Moka",
         product_category == "Electronics")
View(focus_data)

#Checking if the filter by MOKA and product category has worked
unique(focus_data$district)
unique(focus_data$product_category)
nrow(focus_data)

colSums(is.na(focus_data)) %>% sort(decreasing = TRUE)

###class for date/datetime
focus_data <- focus_data %>%
  mutate(date = as.Date(date))

focus_data <- focus_data %>%
  mutate(datetime = ymd_hms(datetime))

###classification of variables
focus_data$channel<-as.factor(focus_data$channel)
focus_data$customer_gender<-as.factor(focus_data$customer_gender)
focus_data$payment_method<-as.factor(focus_data$payment_method)
focus_data$product_subcategory<-as.factor(focus_data
$product_subcategory)
focus_data$customer_segment<-as.factor(focus_data$customer_segment)
focus_data$age_band<-as.factor(focus_data$age_band)
focus_data$product_category<-as.factor(focus_data$product_category)
focus_data$campaign<-as.factor(focus_data$campaign) 

```
The initial dataset had 24000 observations; after applying all the cleaning measures we were left with 23511 observations.
After filtering for Moka, with Electronics as Product category we have 423 observations to work with.

\newpage
**TASK 2 — UNIVARIATE EDA + VISUAL DESIGN**


**2.1 Overview**

This univariate exploratory data analysis (EDA) examines the statistical characteristics of the dataset focusing on the Moka district and the Electronics product category. The purpose of this analysis is to understand
the distribution, variability and statistical behaviour of the variables before conducting further modelling.
The dataset contains 423 observations. The main numerical variables analysed include Net Sales, Gross
Profit, Delivery Minutes and Satisfaction Score. The categorical variables include Payment Method, Customer Gender, Customer Segment and Channel.

````{R}
str(focus_data)
````
The structure output shows the dataset contains both numerical and categorical variables. Missing values
are mainly present Delivery Minutes and Satisfaction Score, which likely reflects unrecorded information
rather than data errors.
\newpage
**2.2 Numeric Distributions**

This section examines the distributional characteristics of the main numeric variables using histograms, density plots, skewness, and kurtosis. The objective is to assess normality, detect skewness, and determine whether transformations are necessary for modelling.

**2.2.1 Distribution of Net Sales**

A histogram combined with a density plot is used to examine the distribution of Net Sales in order to
understand its shape, spread, and possible deviations from normality.
````{R}
ggplot(focus_data,aes(x=net_sales_mur))+
geom_histogram(aes(y=after_stat(density)),
bins=40,fill="steelblue",color="white",na.rm=TRUE)+
geom_density(
color="darkblue",
na.rm=TRUE)+
theme_minimal()+
labs(title="Distribution of Net Sales",x="NetSales(MUR)",y="Density")
````

````{R}
skewness(focus_data$net_sales_mur,na.rm=TRUE)
kurtosis(focus_data$net_sales_mur,na.rm=TRUE)
````
The distribution of Net Sales is approximately symmetric with a slight concentration around the central values. The histogram shows no extreme distortion, and the density curve confirms a fairly balanced shape.  
Skewness is close to zero, indicating minimal asymmetry in the data. This suggests that sales values are relatively stable across transactions, with no strong dominance of extreme high or low values.  
Overall, Net Sales can be considered well-behaved for exploratory analysis, with no need for transformation at this stage.  


**2.2.2 Distribution of Gross Profit**  
The distribution of Gross Profit is analysed to identify possible skewness and variability in profit values.

````{R}
ggplot(focus_data,aes(x=gross_profit_mur))+
geom_histogram(aes(y=after_stat(density)),
bins=40,fill="mediumpurple",color="white",na.rm=TRUE)+
geom_density(color="purple",na.rm=TRUE)+
theme_minimal()
````
````{R}
skewness(focus_data$gross_profit_mur,na.rm=TRUE)
kurtosis(focus_data$gross_profit_mur,na.rm=TRUE)
````
Gross Profit shows moderate positive skewness, with most values clustered in the mid-range and a smooth
right tail reflecting higher-value transactions. The distribution is unimodal and well-structured, with no
significant outliers. Overall, it remains suitable for interpretation in its original scale for univariate analysis.  


**2.2.3 Delivery Minutes Distribution**  
Delivery time is analysed to better understand operational efficiency and variability in service performance.

````{R}
ggplot(focus_data,aes(x=delivery_minutes))+
geom_histogram(aes(y=after_stat(density)),
bins=40,
fill="turquoise",
color="white",
na.rm=TRUE)+
geom_density(
color="darkcyan",
na.rm=TRUE)+
theme_minimal()

cat("Delivery Minutes Skewness:", skewness(focus_data$delivery_minutes, na.rm = TRUE), "\n")
cat("Delivery Minutes Kurtosis:", kurtosis(focus_data$delivery_minutes, na.rm = TRUE), "\n")
````
The distribution of delivery time is moderately positively skewed, with most values clustered between 20
and 40 minutes and a gradual right tail toward longer durations. However, the skewness is not severe, and
the distribution remains smooth and unimodal without extreme outliers. Therefore, a log transformation is
not necessary, as the data is sufficiently well-behaved for analysis in its current form.

**2.2.4 Satisfaction Score Distribution**  
Customer Satisfaction Scores are analysed to understand overall feedback patterns and detect possible response bias.  
````{R}
ggplot(focus_data,aes(x=satisfaction_score))+
geom_histogram(aes(y=after_stat(density)),
bins=20,
fill="orange",
color="white",
na.rm=TRUE)+
geom_density(
color="darkorange",
na.rm=TRUE)+
theme_minimal()

cat("Satisfaction Score Skewness:", skewness(focus_data$satisfaction_score, na.rm = TRUE), "\n")
cat("Satisfaction Score Kurtosis:", kurtosis(focus_data$satisfaction_score, na.rm = TRUE), "\n")
````
Satisfaction Score is nearly perfectly symmetric with no extreme values. The skewness and kurtosis further
confirm the absence of extreme response patterns, so no transformation is required.

**2.3 Outlier Analysis**  
Boxplots are used to detect potential outliers.
````{R}
# Pivot data to long format
df_long <- focus_data %>%
select(net_sales_mur, gross_profit_mur, delivery_minutes) %>%
pivot_longer(cols = everything(),
names_to = "Variable",
values_to = "Value")
# Create combined boxplot
ggplot(df_long, aes(x = Variable, y = Value, fill = Variable)) +
geom_boxplot(na.rm = TRUE) +
scale_fill_manual(values = c("lightblue", "thistle", "mediumturquoise")) +
labs(title = "Boxplots of Key Numeric Variables",
x = "Variable",
y = "Values") +
theme_minimal(base_size = 14) +
theme(legend.position = "none")
```` 
Outliers in the boxplots are identified as individual points beyond the whiskers, which represent values
more than 1.5 times the interquartile range (IQR) from the box. In this case, delivery_minutes shows a
high outlier above the main distribution, while gross_profit_mur also displays an outlier above its upper
whisker. In contrast, net_sales_mur shows no distinct outlier points, indicating that although it has a
wide spread, its values remain within the expected range.  


**2.4 Categorical Distributions (Bar Charts)**  
The distribution of key categorical variables is analysed using bar charts to identify dominant patterns and
underlying customer behaviour across different dimensions.
```{R}
ggplot(focus_data,aes(x=payment_method))+
geom_bar(fill="coral")+
theme_minimal()
```
Card payments dominate, indicating strong preference for convenient payment options.
```{R}
ggplot(focus_data,aes(x=customer_gender))+
geom_bar(fill="pink")+
theme_minimal()
```
Gender distribution is relatively balanced, with a slight predominance of female customers. This suggests
no major gender disparity in purchasing behaviour.  
```{R}
ggplot(focus_data,aes(x=customer_segment))+
geom_bar(fill="mediumseagreen")+
theme_minimal()
```
Returning customers form the largest segment, followed by new customers, with VIP customers being the
smallest group. This indicates strong customer retention and a stable revenue base. 


**Grouped Distribution(Violin Plots)**  

The following violin plots show the distribution of key numeric variables across different groups, highlighting
patterns, variability, and segment-specific trends. The violin plots are presented below and can also be found
in Appendix 4 of the report.
```{R}
ggplot(focus_data, aes(x = customer_segment, y = net_sales_mur, fill = customer_segment)) +
geom_violin() +
labs(title = "Net Sales by Customer Segment — Moka Electronics",
x = "Customer Segment",
y = "Net Sales (MUR)") +
theme_minimal(base_size = 14) +
theme(legend.position = "none",
axis.text.x = element_text(angle = 45, hjust = 1)) +
scale_fill_manual(values = c("tomato","gold","mediumseagreen","skyblue","orchid"))
```
Net Sales are broadly similar across all segments, with Returning customers showing a slightly higher concentration of peak sales around 7,000 MUR.  

```{R}
ggplot(focus_data, aes(x = customer_segment, y = gross_profit_mur, fill = customer_segment)) +
geom_violin() +
labs(title = "Gross Profit by Customer Segment — Moka Electronics",
x = "Customer Segment",
y = "Gross Profit (MUR)") +
theme_minimal(base_size = 14) +
theme(legend.position = "none",
axis.text.x = element_text(angle = 45, hjust = 1)) +
scale_fill_manual(values = c("darkorange","mediumturquoise","pink","limegreen","violet"))
```
Gross Profit is most consistent for VIP customers,while Returning customers exhibit the highest potential
upside with a significant “long tail” of high-profit transactions.  

```{R}
ggplot(focus_data, aes(x = channel, y = delivery_minutes, fill = channel)) +
geom_violin() +
labs(title = "Delivery Minutes by Channel — Moka Electronics",
x = "Channel",
y = "Delivery Minutes") +
theme_minimal(base_size = 14) +
theme(legend.position = "none") +
scale_fill_manual(values = c("skyblue","orange","green"))
```
Delivery times vary by channel: Online orders are generally faster and concentrated around 25 minutes,
Delivery orders are longer and more variable, while the In-store channel has no recorded delivery minutes,
as these values are missing from the dataset.  
Overall, Net Sales and Gross Profit differ slightly across customer segments, with Returning customers
showing higher peaks and potential profit. Delivery times differ by channel, reflecting operational variability
and highlighting areas for potential efficiency improvements.

**Summary Statistics Table**  
The detailed summary statistics table is presented below and can also be found in Appendix 3 of the report.
```{R}
# Ensure required packages are installed and loaded
packages <- c("tidyverse","moments","knitr","kableExtra")
for(p in packages){
if(!require(p, character.only = TRUE)){
install.packages(p)
library(p, character.only = TRUE)
}
}
# Create summary statistics table
summary_table <- data.frame(
Variable = c("Net Sales","Gross Profit","Delivery Minutes","Satisfaction Score"),
Mean = c(
mean(focus_data$net_sales_mur, na.rm=TRUE),
mean(focus_data$gross_profit_mur, na.rm=TRUE),
mean(focus_data$delivery_minutes, na.rm=TRUE),
mean(focus_data$satisfaction_score, na.rm=TRUE)
),
Median = c(
median(focus_data$net_sales_mur, na.rm=TRUE),
median(focus_data$gross_profit_mur, na.rm=TRUE),
median(focus_data$delivery_minutes, na.rm=TRUE),
median(focus_data$satisfaction_score, na.rm=TRUE)
),
Min = c(
min(focus_data$net_sales_mur, na.rm=TRUE),
min(focus_data$gross_profit_mur, na.rm=TRUE),
min(focus_data$delivery_minutes, na.rm=TRUE),
min(focus_data$satisfaction_score, na.rm=TRUE)
),
Max = c(
max(focus_data$net_sales_mur, na.rm=TRUE),
max(focus_data$gross_profit_mur, na.rm=TRUE),
max(focus_data$delivery_minutes, na.rm=TRUE),
max(focus_data$satisfaction_score, na.rm=TRUE)
),
Skewness = c(
moments::skewness(focus_data$net_sales_mur, na.rm=TRUE),
moments::skewness(focus_data$gross_profit_mur, na.rm=TRUE),
moments::skewness(focus_data$delivery_minutes, na.rm=TRUE),
moments::skewness(focus_data$satisfaction_score, na.rm=TRUE)
),
Kurtosis = c(
moments::kurtosis(focus_data$net_sales_mur, na.rm=TRUE),
moments::kurtosis(focus_data$gross_profit_mur, na.rm=TRUE),
moments::kurtosis(focus_data$delivery_minutes, na.rm=TRUE),
moments::kurtosis(focus_data$satisfaction_score, na.rm=TRUE)
)
)
# Display table
knitr::kable(summary_table,
caption = "Summary statistics of numeric variables (including skewness and kurtosis)",
digits = 2) %>%
kableExtra::kable_styling(full_width = FALSE, position = "center")
```
**Conclusion**  
The univariate analysis shows that the dataset is generally well-structured and suitable for further analysis.
The numerical variables are mostly well-behaved, with Net Sales and Satisfaction Score showing approximately symmetric distributions, while Gross Profit and Delivery Minutes display mild positive skewness due
to higher transaction values and longer delivery times. A small number of outliers are present, mainly in
Gross Profit and Delivery Minutes, reflecting normal operational variation. The categorical analysis shows a
strong preference for card payments, a dominant returning customer base, and a relatively balanced gender
distribution. Differences in Delivery Minutes across channels highlight variation in operational efficiency.
Overall, the dataset is consistent, reliable, and appropriate for further analysis.
\newpage
**Task 3 — Bivariate & Multivariate EDA**  
**3.1 Overview**

This section presents the bivariate and multivariate exploratory data analysis (EDA). The aim is to
understand how different variables are related and how they affect business performance and
customer behaviour.
The analysis focuses on key variables such as net sales, gross profit, customer satisfaction. It also
compares results across stores, sales channels, and customer characteristics like gender and age
group.  

**3.2 Net Sales by Store and Channel**  

The bar chart reveals a significant reliance on physical retail within the IslandMart ecosystem, as in-store sales dramatically outperform both online and delivery channels at Store 2 and Store 11. 
```{r}
focus_data %>%
  group_by(store_id, channel) %>%
  summarise(total_net_sales = sum(net_sales_mur, na.rm = TRUE)) %>%
  ggplot(aes(x = factor(store_id), y = total_net_sales, fill = channel)) +
  geom_col(position = "dodge") +
  labs(title = "Net Sales by Store and Channel",
       x = "Store ID",
       y = "Total Net Sales (MUR)",
       fill = "Channel") +
  theme_minimal()
```
 
**3.3 Net Sales by Gender and Store**  
The bar chart provided offers a detailed demographic breakdown of net sales across IslandMart locations, revealing distinct customer purchasing behaviors between Store 2 and Store 11.

```{r}
focus_data %>%
  group_by(store_id, customer_gender) %>%
  summarise(total_net_sales = sum(net_sales_mur, na.rm = TRUE)) %>%
  ggplot(aes(x = factor(store_id), y = total_net_sales, fill = customer_gender)) +
  geom_col(position = "dodge") +
  labs(title = "Net Sales by Gender and Store",
       x = "Store ID",
       y = "Total Net Sales (MUR)",
       fill = "Gender") +
  theme_minimal()

```

**3.4 Multivariate Analysis (Pair Plot)**  
This correlation matrix provides a sophisticated multivariate analysis of IslandMart’s operational health, specifically isolating the drivers of profitability and customer experience for Store 2 and Store 11. 
```{r echo=FALSE}
vars_num <- c(
  "net_sales_mur",
  "gross_profit_mur",
  "delivery_minutes",
  "discount_pct",
  "quantity",
  "satisfaction_score"
)

p_pairs <- ggpairs(
  focus_data,
  columns = vars_num,
  aes(color = factor(store_id), alpha = 0.3),
  upper = list(continuous = wrap("cor", size = 3)),
  lower = list(continuous = wrap("points", size = 0.7))
) +
  theme_bw() +
  theme(
    legend.position = "bottom",
    panel.border = element_rect(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  ) +
  labs(title = "Multivariate Relationships Between Key Variables")

p_pairs
```

**3.5 Product Subcategory by Age Band per Store**  
This heat map provides a multivariate view of IslandMart’s customer demographics, crossing product subcategories with age bands across Store 2 and Store 11.
```{r}
focus_data %>%
  count(store_id, age_band, product_subcategory) %>%
  ggplot(aes(x = age_band, y = product_subcategory, fill = n)) +
  geom_tile() +
  facet_wrap(~ store_id) +
  labs(title = "Product Subcategory by Age Band per Store",
       x = "Age Band",
       y = "Product Subcategory",
       fill = "Count") +
  theme_minimal()
```
\newpage
**Task 4 : Time-series + special visualization **  
Under this section, we will look at the sales and profit trends of Electronics of Moka District.  

**4.1 Time series Analysis**  
We have two graphs showing the monthly net sales and net sales vs profit respectively. 

```{R}
monthly_sales <- focus_data %>%
  mutate(month = floor_date(date, "month")) %>%
  group_by(month) %>%
  summarise(
    total_sales = sum(net_sales_mur, na.rm = TRUE),
    total_profit = sum(gross_profit_mur, na.rm = TRUE)
  )
ggplot(monthly_sales,aes(x = month, y = total_sales)) +
  geom_line(color = "lightblue", linewidth=1) +
  labs(title = "Monthly Net Sales Trend",
       x = "Month",
       y = "Total Net Sales") +
theme_minimal()
```
The monthly sales trend shows whether Electronics sales in Moka are stable, growing or fluctuating over time.

```{R}
ggplot(monthly_sales, aes(month)) +
  geom_line(aes(y = total_sales, color = "Sales")) +
  geom_line(aes(y = total_profit, color = "Profit")) +
  labs(title = "Monthly Sales vs Profit",
       y = "Amount",
       color = "Metric") +
  theme_minimal()
```
Comparing sales and profit helps determine whether higher sales are consistently translating into higher profitability.  

**4.2 Store Performance and Spacial Analysis**  
To compare Moka with the wide store network, store-level performance is illustrated below.

```{R}
store_map <- data %>%
  group_by(store_name, district, store_lat, store_lon) %>%
  summarise(total_sales = sum(net_sales_mur),
            total_profit = sum(gross_profit_mur),
            .groups = "drop") %>%
  filter(!is.na(store_lat), !is.na(store_lon))

ggplot(store_map, aes(x = store_lon, y = store_lat)) +
  geom_point(aes(size = total_sales, color = district), alpha = 0.7) +
  labs(title = "Store Performance by Location",
       x = "Longitude",
       y = "Latitude",
       size = "Total Sales") +
  theme_minimal()
```

```{R}
ggplot(store_map, aes(store_lon, store_lat)) +
  geom_point(aes(size = total_sales,
                 color = district == "Moka"), alpha = 0.8) +
  scale_color_manual(values = c("TRUE" = "red", "FALSE" = "grey"),
                     guide = "none") +
  labs(title = "Store Sales Map (Moka)",
       x = "Longitude",
       y = "Latitude",
       size = "Sales") +
  theme_minimal()
```
The Moka highlighted map shows where Moka sits spatially relative to the broader store network.

\newpage
The following table and chart are found in **Appendix 2**

**Appendix proof- Districts and No.of stores**
```{R}
data_summary <- data %>%
  group_by(district) %>%
  summarise(
    number_of_stores = n_distinct(store_name),
    number_of_transactions = n()
  ) %>%
  arrange(desc(number_of_transactions))

print(data_summary)
```
**Transactions count by Districts**
```{R}
data %>%
  count(district) %>%
  ggplot(aes(x = reorder(district, n), y = n)) +
  geom_bar(stat = "identity", fill = "mediumblue") +
  geom_text(aes(label = n), hjust = -0.1) +
  coord_flip() +
  labs(title = "Transaction Volume by District",
       x = "District",
       y = "Number of Transactions") +
  theme_minimal()
```

**Moka Monthly Transactions**
```{R}
focus_data %>%
  count(month) %>%
  mutate(month_name = month.name[month]) %>%
  ggplot(aes(x = factor(month_name, levels = month.name), y = n)) +
  geom_bar(stat = "identity", fill = "lightblue") + 
  geom_text(aes(label = n), vjust = -0.3, size = 4) +
  labs(title = "Number of Transactions by Month (Moka – Electronics)",
       x = "Month",
       y = "Transactions") +
  theme_minimal()
```
